Ae. albopictus CPP manuscript

### 2020-11-11 Albopictus_cpp_MS.Rmd
added alt deadline data

```{r eval=FALSE, message=FALSE, include=FALSE}
install.packages("tidyverse") #1.2.1
install.packages("drc") #3.0.1
install.packages("stringr")
install.packages("lubridate")
install.packages(c("maps", "mapdata"))
devtools::install_github("dkahle/ggmap")
install.packages("scatterpie")
install.packages("ggedit")
install.packages("ggrepel")
install.packages("papeR")
install.packages("cowplot")
install.packages("flextable")
install.packages("stats")
install.packages("ggbiplot")
```

```{r message=FALSE, echo=FALSE, include=FALSE}
library("papeR")
library("drc") #3.0.1
library("knitr") #1.25
library("tidyverse") #1.2.1
library("stringr") #1.4.0
library("geosphere") #1.5.10
library("lubridate") #1.7.4
library("maps")
library("mapdata")
library("ggedit")
library("ggrepel")
library("cowplot")
library("ggrepel")
library("stats")
library("ggbiplot")

@@

```{r collate_cpp_values, echo=FALSE, include=FALSE}
cpp_2018<-data.frame(ED(d.model.fit, 50))
cpp_2018<-rownames_to_column(cpp_2018, var="pop") 
cpp_2018$pop<-str_sub(cpp_2018$pop,3,5)
popdata_2018<-DI_summary %>% group_by(pop) %>% 
  dplyr::summarize(lat=first(lat), country=first(country))
cpp_2018<-right_join(cpp_2018, popdata_2018, by="pop")
cpp_2018$year<-rep("2018", nrow(cpp_2018))
colnames(cpp_2018)[colnames(cpp_2018)=="Std..Error"] <- "StdError"
cpp_2008<-data.frame(ED(d.model.fit.2008, 50)) %>% rownames_to_column(var="pop")
cpp_2008$pop<-str_sub(cpp_2008$pop,3,5)
colnames(cpp_2008)[colnames(cpp_2008)=="Std..Error"] <- "StdError"
popdata_2008<-DI_summary_2008 %>% group_by(pop) %>% 
  dplyr::summarize(country=first(country), lat=first(lat), year=first(year)) 
cpp_2008<-right_join(cpp_2008, popdata_2008, by="pop")
#1988 data not used in present study but can load it anyway
cpp_1988<-data.frame(read.csv("CPP_1988.csv")) 
cpp_1988$pop<-na_if(cpp_1988$pop, "NA")
cpp_1988$StdError<-na_if(cpp_1988$StdError, "NA")

@@

```{r, echo=FALSE, include=FALSE}
#basic breakdown
ggplot(cpp_all, aes(x=lat, y=Estimate, color=country)) + 
  geom_point() + 
  geom_errorbar(aes(ymin=Estimate-StdError, ymax=Estimate+StdError), na.rm=TRUE) + 
  facet_grid(~year) + stat_smooth(method=lm, se=FALSE)
```

@@

```{r maps, echo=FALSE, include=FALSE}
usa<-map_data("usa")
jp<-map_data("world") %>% filter(region=="Japan")
pops<-data.frame(read.csv("mapping_pops_cpp_expt.csv"))
us_map<-ggplot() + geom_polygon(data = usa, aes(x=long, y = lat, group=group), fill=NA, color="blue") + 
  coord_fixed(1.3) + geom_point(data = filter(pops, country=="US"), aes(x = longitude, y = latitude), size = 1) + xlim(c(-85, -65)) +
  geom_label(data=filter(pops, country=="US"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines")) +
  theme_bw() 
japan_map<-ggplot() + geom_polygon(data = jp, aes(x=long, y = lat, group=group), fill=NA, color="red") + 
  coord_fixed(1.3) + geom_point(data = filter(pops, country=="JP", only_2008=="n"), aes(x = longitude, y = latitude), size = 1)  + xlim(c(125, 147)) +
  geom_label_repel(data=filter(pops, country=="JP", only_2008=="n"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines")) +
  theme_bw() + theme(axis.title.x=element_blank(), axis.title.y=element_blank())

#bw locations map
us_map_bw<-ggplot() + geom_polygon(data = usa, aes(x=long, y = lat, group=group), fill=NA, color="dark gray") + 
  coord_fixed(1.3) + geom_point(data = filter(pops, country=="US"), aes(x = longitude, y = latitude), size = 2, shape=21, color="black", fill="gray", stroke=1) + xlim(c(-87, -65)) + ylim(c(24,48)) +
  geom_text_repel(data=filter(pops, country=="US"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines"), nudge_x=1.5, size=6) + theme_bw() + theme(text=element_text(size=24), plot.margin = unit(c(0,0,0,0), "cm"), panel.grid=element_line(size=0.1)) + xlab("") + ylab("")

@@

gdd_08_sameyear <- ggplot(filter(gdd465_all_new, year==2008, CPP_YEAR==2008), aes(x=gdd465, y=Estimate, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=4, stroke=1, shape=22) + stat_smooth(method=lm, se=FALSE, aes(color=country), linetype="longdash") + scale_color_manual(values=c("black","gray"), name="Country") +
  theme(legend.position=c(.25,.25), text=element_text(size=22),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
    xlab("GDD Deadline 2008 \n(ordinal day)") + ylab("CPP (h), 2008") + ylim(12.3,14.2)

@@

gdd_18_sameyear <- ggplot(filter(gdd465_all_new, year==2018, CPP_YEAR==2018), aes(x=gdd465, y=Estimate, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=4, stroke=1, shape=21) + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black","gray"), name="Country") +
  theme(legend.position=c(.25,.25), text=element_text(size=22),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
xlab("GDD Deadline 2018 (ordinal day)") + ylab("CPP (h), 2018") + ylim(12.3, 14.2)

@@

```{r}
#Fig S4 two parts
gdd_examples<-plot_grid(gdd_deadline_plots, sample_locations_gdd, nrow=2, labels="AUTO")

>>>>

```{r eval=FALSE, message=FALSE, include=FALSE}
install.packages("tidyverse") #1.2.1
install.packages("drc") #3.0.1
install.packages("stringr")
install.packages("lubridate")
install.packages(c("maps", "mapdata"))
devtools::install_github("dkahle/ggmap")
install.packages("ggedit")
install.packages("ggrepel")
install.packages("papeR")
install.packages("cowplot")
install.packages("flextable")
install.packages("stats")
install.packages("ggbiplot")
```

```{r message=FALSE, echo=FALSE, include=FALSE}
library("papeR")
library("knitr") 
library("drc") #3.0.1
library("tidyverse") #1.3.0
library("stringr") #1.4.0
library("lubridate") #1.7.4
library("maps") #3.3.0
library("mapdata") #2.3.0
library("cowplot") #1.0.0
library("stats") #3.6.2
library("ggbiplot") #0.55
library("ggrepel")
```

@@

```{r collate_cpp_values, echo=FALSE, include=FALSE}
cpp_2018<-data.frame(ED(d.model.fit, 50))
cpp_2018<-rownames_to_column(cpp_2018, var="pop") 
cpp_2018$pop<-str_sub(cpp_2018$pop,3,5)
popdata_2018<-DI_summary %>% group_by(pop) %>% 
  dplyr::summarize(lat=first(lat), country=first(country))
cpp_2018<-right_join(cpp_2018, popdata_2018, by="pop")
cpp_2018$year<-rep("2018", nrow(cpp_2018))
colnames(cpp_2018)[colnames(cpp_2018)=="Std..Error"] <- "StdError"
cpp_2008<-data.frame(ED(d.model.fit.2008, 50)) %>% rownames_to_column(var="pop")
cpp_2008$pop<-str_sub(cpp_2008$pop,3,5)
colnames(cpp_2008)[colnames(cpp_2008)=="Std..Error"] <- "StdError"
popdata_2008<-DI_summary_2008 %>% group_by(pop) %>% 
  dplyr::summarize(country=first(country), lat=first(lat), year=first(year)) 
cpp_2008<-right_join(cpp_2008, popdata_2008, by="pop")

@@

@@

```{r maps, echo=FALSE, include=FALSE}
usa<-map_data("usa")
jp<-map_data("world") %>% filter(region=="Japan")
pops<-data.frame(read.csv("mapping_pops_cpp_expt.csv"))

us_map_bw<-ggplot() + geom_polygon(data = usa, aes(x=long, y = lat, group=group), fill=NA, color="dark gray") + 
  coord_fixed(1.3) + geom_point(data = filter(pops, country=="US"), aes(x = longitude, y = latitude), size = 2, shape=21, color="black", fill="gray", stroke=1) + xlim(c(-87, -65)) + ylim(c(24,48)) +
  geom_text_repel(data=filter(pops, country=="US"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines"), nudge_x=1.5, size=6) + theme_bw() + theme(text=element_text(size=24), plot.margin = unit(c(0,0,0,0), "cm"), panel.grid=element_line(size=0.1)) + xlab("") + ylab("")

@@

@@

@@

```{r}
#Fig S4 two parts
gdd_examples<-plot_grid(gdd_deadline_plots, sample_locations_gdd, nrow=2, labels="AUTO")
```

```{r}
#range of possible deadlines
alt_deadlines<-read.csv("gdd_alternate_deadlines.csv")
#France location in 2009
filter(alt_deadlines, Country=="FRA", Year==2009)
```

``` {r}
#range of possible deadlines w/latitude/CPP info
deadlines_range<-left_join(gdd_all_new, alt_deadlines, by=c("pop"="Location", "year"="Year")) %>% select(pop, year, latitude, country, only_2008, Estimate, CPP_YEAR, gdd465, Deadline_GDD_300, Deadline_GDD_350, Deadline_GDD_400, Deadline_GDD_450, Deadline_GDD_500, Deadline_GDD_550, Deadline_GDD_600)

ggplot(filter(deadlines_range, year==2017, CPP_YEAR==2018), aes(x=latitude, y=Deadline_GDD_600, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=2, stroke=1, shape=21) + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black","gray"), name="Country") +
  theme(legend.position=c(.25,.25), text=element_text(size=22),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
xlab("Latitude") + ylab("Deadline")

data_18_range<-deadlines_range %>% filter(CPP_YEAR==2018) %>% filter(year==2017)
model300<-lm(Estimate~Deadline_GDD_300*country, data=data_18_range)
model350<-lm(Estimate~Deadline_GDD_350*country, data=data_18_range)
model400<-lm(Estimate~Deadline_GDD_400*country, data=data_18_range)
model450<-lm(Estimate~Deadline_GDD_450*country, data=data_18_range)
model500<-lm(Estimate~Deadline_GDD_500*country, data=data_18_range)
model550<-lm(Estimate~Deadline_GDD_550*country, data=data_18_range)
model600<-lm(Estimate~Deadline_GDD_600*country, data=data_18_range)

rsq<-c(summary(model300)$r.squared, summary(model300)$r.squared, summary(model350)$r.squared, summary(model400)$r.squared, summary(model450)$r.squared, summary(model500)$r.squared, summary(model550)$r.squared, summary(model600)$r.squared)

data_18_range_long<-pivot_longer(data_18_range, Deadline_GDD_300:Deadline_GDD_600, names_to="deadlinetype", values_to="deadline")

ggplot(filter(data_18_range_long, year==2017, CPP_YEAR==2018), aes(x=latitude, y=deadline, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=2, stroke=1, shape=21) + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black","gray"), name="Country") +
  theme(legend.position="bottom",
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
xlab("Latitude") + ylab("Deadline") + facet_wrap(~deadlinetype)

```{r}

#add range of possible deadline values to Manassas 2014
man_deadlines_range<-filter(deadlines_range, pop=="MAN", year==2014, CPP_YEAR==2018) %>% select(Deadline_GDD_300:Deadline_GDD_600)  %>% t() %>% data.frame() %>% pull()

man2014_gdd_alt_plot <-
  ggplot(filter(man2014, DOY_2014 %in% c(238:278)), aes(x=DOY_2014, y=degrees, fill=degreetype)) +
    theme(text=element_text(size=24), legend.position="none",
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
  geom_bar(stat="identity", width=1) + 
  scale_fill_manual(values=c("dark gray","light gray")) + 
  geom_hline(yintercept=10, linetype="dashed") + geom_vline(xintercept=251) + geom_vline(xintercept=man_deadlines_range, linetype="dotted") + 
  annotate("text", x=(man_deadlines_range)-0.5, y=c(1,3,5,7,9,11,13), label=c("GDD300", "GDD350", "GDD400","GDD450","GDD500","GDD550","GDD600"),angle=90) +
  geom_vline(xintercept=252.5, color="red") + annotate("text",x=250.5, y=20, label="GDD-465 (model used)", fontface="bold", angle=90) + annotate("text",x=253, y=20, label="50% diapause in the field", color="red", angle=90) +
  xlab("Ordinal Day (2014)") + ylab("Mean Daily Temp (C)")
